```{r include = FALSE}
library(DataComputing)
show_answers <- TRUE
```

The following tasks can be performed using a single data verb. For each task, pick out the appropriate verb from this set: `summarise()`, `mutate()`, `arrange()`, `filter()`, `select()`, `group_by()`.


a. Find the average of one of the variables.
b. Add a new column that is the ratio between two variables.
c. Sort the cases in descending order of a variable.
d. Create a new data table that includes only those cases that meet a criterion.
e. From a data table with several variables produce an output that has the same cases but only the first two variables.
f. From a data table with a categorical variables and a quantitative variable produce a count for the number of cases for each level of the categorical variable.


```{block echo = show_answers}
**Solution:**

We'll use the `mtcars` data table to show the answers, although of course any data table will do. Here are a few cases from `mtcars`.
```
```{r echo = FALSE, eval = show_answers}
set.seed(101)
mtcars %>% sample_n(3) %>% knitr::kable()
```



```{r  echo = show_answers, eval = show_answers, results = "hide"}
# a.
mtcars %>%
  summarise(m_cyl=mean(cyl))
```

```{r echo = show_answers, eval = show_answers, results = "hide"}
# b.
mtcars %>%
  mutate(ratio=mpg/cyl) 
```

```{r  echo = show_answers, eval = show_answers, results = "hide"}
# c.
mtcars %>%
  arrange(desc(mpg)) 
```

```{r echo = show_answers, eval = show_answers, results = "hide"}
# d.
mtcars %>%
  filter(mpg >30)
```

```{r echo = show_answers, eval = show_answers, results = "hide"}
# e. 
mtcars %>%
  select(mpg, cyl)
```

```{r echo = show_answers, eval = show_answers, results = "hide"}
# f.
mtcars <- mtcars %>%
  select(mpg, cyl, disp)
```

Credit: Adam Lucas
