---
output: html_notebook
---

# More transformation verbs

```{r include = FALSE}
library(DataComputing)
```

Recall that a *transformation verb* takes a variable as an input and produces a variable of the same length as the output. 

Some familiar transformation verbs from primary and secondary school:

- arithmetic operations (`+`, `-`, `/`, `*`) 
- mathematical functions such as logs and exponentiation

There are additional transformations which will be unfamiliar, simply because they did not fit into the algebra- and trigonometry-based high-school curriculum

Two of these transformation operations are

1. Rank transforms.
2. Lead and lag transforms.

## Rank transforms 

Many questions take forms such as these:

* "Find the largest ..."
* "Find the three largest ..."
* "Find the smallest within each group ..."

The functions `min()` and `max()` are *reduction* verbs. They tell you the single lowest or highest value in a set. Because they are reduction verbs, they are often used in `summarise()`, which reduces a set of cases to a single case.

```{r}
BabyNames %>% 
  group_by(sex) %>%
  summarise(most_popular = max(count))
```

Notice that `name` was not carried along. When you `summarise()`, the only variables that appear in the output are the grouping variables, and the variables you create through the arguments to `summarise()`.

If you wanted to know the *names* that are most popular, you will need to rely on `filter()`. Get rid of the cases that are not the most popular.

```{r}
BabyNames %>% 
  group_by(sex) %>%
  filter(count == max(count))
```


`Filter()` needs a criterion.  The criterion `count == max( count )` (with the double equals sign `==`) passes through the case where the value of `count`  matches the largest value of `count`.  That will be the biggest case.

```{r}
BabyNames %>% 
  group_by(sex) %>%
  filter(count > 0.90 * max(count))
```

Note: Almost all of these are from the late 1940s and early 1950s. In part, this reflects the baby boom. Perhaps it also reflects the conformity that people associate with that era.

QUESTION: How would you estimate "conformity" for each year?

Possibility: Take the 10 most popular names each year. Find out what fraction of the total number of births that was.

The `rank()` operation is helpful here.

The `rank()` function does something simple but powerful: it replaces each number in a set with where that number stands with respect to the others.  For instance, look at the tiny data table `Set` shown in Table \ref{tab:set-of-numbers}. What's the rank of the number 5 in the `numbers` variable.
```{r}
Set <- data.frame(numbers = as.integer(c(2, 5, 4, 7, 2, 9, 9, 8)))
Set %>%
  mutate(the_rank = rank(numbers))
```
Or, seen another way
```{r}
Set %>%
  mutate(the_rank = rank(numbers)) %>%
  arrange(numbers)
```

Notice how ties are broken. Also note that the biggest numbers have the *highest* ranks. This is different than the convention in everyday language, where the "Number 1 ranked team" is the best team.  To follow this convention, use `rank(desc(numbers))`.

```{r}
Set %>%
  mutate(the_rank = rank(desc(numbers))) %>%
  arrange(numbers)
```

Suppose you want to find the 3rd most popular name of all time. Use `rank()`. 
```{r}
BabyNames %>% 
  group_by(name) %>%
  summarise(total = sum(count)) %>%
  filter(rank(desc(total)) == 3) 
```

Or, to find the top three most popular names, replace `==` in the above by `<=`.

```{r}
BabyNames %>% 
  group_by(name) %>%
  summarise(total=sum(count)) %>%
  filter( rank( desc(total) ) <= 3) 
```

When applied to grouped data, `rank()` will be calculated separately within each group.  That is, the rank of a value will be with respect to the other cases in that group.  For instance, here's the third most popular name for each sex. 

```{r}
BabyNames %>% 
  group_by(sex) %>% 
  filter(rank(desc(count)) == 3) 
```

### Tied ranks

Sometimes, two or more numbers are tied in rank. The `rank()` function deals with these by assigning all the tied values the same rank, which is the mean of the ranks those values would have had if they were even slightly different. There are other rank-like transformation verbs that handle ties differently. For instance, `row_number()` breaks ties in favor of the first case encountered.

```{r}
Set %>% 
  mutate(the_rank = rank(numbers), 
         ties_broken = row_number(numbers)) %>%
  arrange(numbers)
```

### Example: Conformity

Calculate the total number of babies given the 10 most popular names each year as a fraction of the total number of babies born that year.


#### A two-table method
```{r}
Popular <- 
  BabyNames %>%
  group_by(year) %>%
  filter(row_number(desc(count)) <= 10) %>%
  summarise(total_popular = sum(count)) 
AllNames <-
  BabyNames %>% 
  group_by(year) %>%
  summarise(total = sum(count))
```
 
```{r}
head(Popular)
head(AllNames)
```

Question: What would have happened if `group_by()` came after `filter()`?

We'll need an operation to combine two tables to find the proportion each year. Here's a teaser:
```{r}
AllNames %>%
  left_join(Popular) %>%
  mutate(pop_prop = total_popular / total) %>%
  ggplot(aes(x = year, y = pop_prop)) + 
  geom_line() + 
  ylab("Conformity Index")
```


## Leads and lags

Find the names that increase the most from one year to the next. Let's take "increase the most" to mean "the biggest proportional increase", but consider only names that have more than 100 kids in the earlier year.

Find proportional increase for each name for each year, but push this to zero if the base year had less than 100 kids.

```{r}
Sharp_increases <-
  BabyNames %>% 
  group_by(name, year) %>%
  summarise(births = sum(count)) %>%
  arrange(year) %>%
  mutate(change = (births > 100) * lead(births, 1) / births) %>%
  ungroup() %>%
  filter(row_number(desc(change)) <= 100) %>%
  arrange(desc(change)) 
```

Notice "Woodrow" and "Wilson" in 1911 to 1912. Why?

Notice "Samantha" and "Darin" in 1964. Why? (Hint: "Bewitched")


FLAW: There might be years left out for some names. We'll have to wait until we study joins to see how to fix this.
