```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(printr)
library(mosaic)
knitr::opts_chunk$set(tidy=FALSE, message=FALSE)
options(width = 80)
```


# Combining data from different sources

Glyph-ready data often combines data from different sources.

* Perhaps they come from different experiments or institutions.
* Often, they were collected with different objectives than yours.
* Perhaps they are completely different types of data, 



*Example: Medicare data*

* `MedicareProviders`: Name and location
* `DirectRecoveryGroups`: Descriptions of standardized medical procedures
* `MedicareCharges`: Charges and payments for different DRGs by different providers
* `ZipDemographics`: Population and age structure in each ZIP code.

 

**Example: Holiday births**

**Goal**: Compare the number of births on holidays to those on non-holiday weekdays.

* The `mosaicData::Birthdays` table gives the daily number of births in US states from 1969 to 1988.
* `Birthdays` doesn't *directly* tell us which days are holidays, but ...

We can use the `date` variable to look up each case in a list of holidays, e.g. 

```{r echo=1, cache = TRUE}
Holidays <- read.file("http://tiny.cc/dcf/US-Holidays.csv")
set.seed(101)
Holidays %>%
  sample_n(size=4) %>%
  arrange(year)
```


## Relational databases

Storing data in separate tables can be beneficial even when the data are coming from the same source: 

* There is no "one size fits all" glyph-ready format.  Often the kinds of analysis that will be done are not specifically anticipated when data are collected. 
* Glyph-ready data often contains **redundancies**.  This makes it hard to update or correct data.

**Strategy**: Don't even try to smash all data into one big table.  Instead, join related tables as needed.

 

**Example: Grades and Enrollment**

```{r}
Grades <- read.file("http://tiny.cc/mosaic/grades.csv")
```
```{r echo=FALSE}
set.seed(101)
Grades %>% sample_n(size=4)
```

```{r}
Courses <- read.file("http://tiny.cc/mosaic/courses.csv")
```
```{r echo=FALSE}
set.seed(101)
Courses %>% sample_n(size=3)
```

 

## Joins

A *join* is a data verb that combines two tables. 

- These care called the *left* and the *right* tables.

There are several kinds of join.

* All involve establishing a correspondance --- a match --- between each case in the left table and zero or more cases in the right table.
* The various joins differ in how they handle multiple matches or missing matches.

    

## Example: Average class size

**Goal**: Figure out the average class size seen by each student.

* `enroll` comes from `Courses` table.
* Student (`sid`) comes from `Grades`.
* `sessionID` is in both tables.

```{r}
Grades %>%
  left_join(Courses) %>% sample_n(size=4)
```


Once `Courses` and `Grades` are joined, it's straightforward to find the average enrollment seen by each student.

```{r}
AveClassEachStudent <- Grades %>% 
  left_join(Courses) %>%
  group_by(sid) %>%
  summarise(ave_enroll = mean(enroll, na.rm=TRUE))
```
```{r echo=FALSE}
AveClassEachStudent %>% sample_n(3) # look at just a few
```

   -

**Statistical Digression**

Why are these numbers different?

```{r}
AveClassEachStudent %>% summarise(average = mean(ave_enroll))
Courses %>% summarise(average = mean(enroll))
```
 
  

## Establishing a match between cases

A match between a case in the *left* table and a case in the *right* table is made based on the values in pairs of corresponding variables.

* **You** specify which pairs to use.
* A pair is a variable from the left table and a variable from the right table.
* Cases must have *exactly equal* values in the left variable and right variable for a match to be made.

**Example**:
```{r echo=1:2}
Grades %>% 
  left_join(Courses, by = c(sessionID = "sessionID")) %>%
  head(4)
```

The default value of `by=` is all variables with the same names in both tables.     
- This is **not reliable** unless you've checked.

 

## Kinds of join

Different kinds of join have different answers to these questions.

* What to do when there is **no match** between a left case and any right case?
* What to do when there are **multiple matching cases** in the right table for a case in the left table?

Most popular joins: `left_join()` and `inner_join()`


1. No match of a left case to a right case
    * `left_join()` Keep the left case and fill in the new variables with NA
    * `inner_join()` Discard the left case.
    Less popular joins:
    * `full_join()` Keep left case as well as unmatched right cases.
    * `semi_join()` Discard the left case.
    * `anti_join()` Keep the left case but discard any left case with a match in the right table
2. Multiple matches of right cases to a left case
    * `left_join()` and `inner_join()` do the same thing: Keep **all combinations**.
    Less popular joins:
    * `full_join()` Keep all combinations.
    * `semi_join()` Keep just one copy of the left case.
    * `anti_join()` Discard the left case.

## Example: Grade-point averages

Here are three data tables relating student grades in courses at Macalester in 2005

```{r echo=c(1,3,5)}
Grades <- read.file("http://tiny.cc/mosaic/grades.csv")
Grades %>% head(1)
Courses <- read.file("http://tiny.cc/mosaic/courses.csv")
Courses %>% head(1)
GradePoint <- read.file("http://tiny.cc/mosaic/grade-to-number.csv")
GradePoint %>% head(1)
```

 

## Activity: Which to Join?

For each of these, say what tables you would need to join and what the corresponding variables will be.

1. How many students in each department?
#. What fraction of grades are below B+ in each department?
#. What's the grade-point average (GPA) for each student?
#. Grade-point average for each department or instructor
#. What's the 95% confidence interval on the GPA for each student?
#. (Statistically more sophisticated) To what extent does the grade reflect the student or the department or instructor?

 

## Example: Joining for cleaning

Birds at the Ordway Nature Preserve

```{r eval=FALSE}
devtools::install_github("DataComputing/DataComputing")
require(DataComputing)
View(OrdwayBirds)
help(OrdwayBirds)
```

As a group, let's correct the species names using [this data table](https://docs.google.com/spreadsheets/d/1GDKeulVSRXL7Q_5Q0a7Offc_IPX2T4sZ80VN5VLASew/edit#gid=0).

* Count how many birds there are of each species in `OrdwayBirds` using the corrected species names.


