# Regex examples.

```{r include = FALSE}
library(DataComputing)
```

Here are some examples of patterns in names and the use of a regular expression to detect them. We'll work with the baby names data, summarised to give the total count of each name for each sex.

```{r}
NameList <-
  BabyNames %>%
  mutate(name = tolower(name)) %>%
  group_by(name, sex) %>%
  summarise(total=sum(count)) %>%
  arrange(desc(total))
```



The regular expression is the string in quotes.  `grepl()` is a function that compares a regular expression to a string, returning TRUE if there's a match, FALSE otherwise.

1. The name contains "shine", as in "sunshine" or "moonshine"
    ```{r}
NameList %>%
  filter(grepl("shine", name)) %>%
  head()
```
2. The name contains three or more vowels in a row.
```{r}
NameList %>%
  filter(grepl("[aeiou]{3,}", name)) %>%
  head()
```
3. The name contains three or more consonants in a row.
```{r}
NameList %>%
  filter(grepl("[^aeiou]{3,}", name)) %>%
  head()
```
4. The name contains "mn"
```{r}
NameList %>%
  filter(grepl( "mn", name)) %>%
  head()
```

5. The first, third, and fifth letters are consonants.
```{r}
NameList %>%
  filter(grepl("^[^aeiou].[^aeiou].[^aeiou]", name)) %>%
  head()
```
6. How often do boys' and girls' names end in vowels?

```{r}
NameList %>%
  filter( grepl( "[aeiou]$", name ) ) %>%
  group_by( sex ) %>%
  summarise( total=sum(total) )
```

Girls' names are almost five times as likely to end in vowels as boys' names.

7. What are the most common end vowels for names?

To answer this question, you have to extract the last vowel from the name.  The `extractMatches()` transformation function can do this.

```{r}
NameList %>%
  extractMatches( "([aeiou])$", name, vowel=1 ) %>%
  group_by( sex, vowel ) %>%
  summarise( total=sum(total) ) %>%
  arrange( sex, desc(total) )
```

Credit: Adam Lucas
