# Example: Finding the PO number in an address

```{r include = FALSE}
library(DataComputing)
```

Suppose you wanted to extract the PO Box number from an address.

Read the street address data and pull out a sample of a few dozen cases.
```{r cache=TRUE}
Addresses <- read.file("http://tiny.cc/dcf/street-addresses.csv")
Sample <- Addresses %>%
  sample_n(size = 50)
```

Following each of the steps listed above:

1. The PO Box cases tend to have a substring "PO".
2. The regular expression for "PO" is simply `"PO"`.^[Perhaps you want to insist that PO be preceeded and followed by a space or the start of the address, so that the regex doesn't match, say, `"49 EDGAR ALLEN POE TERRACE"` or `"POPCORN DRIVE"`. This regex would look like `"^|\\s+PO\\s+"`. But in the example, we'll keep it simple.]
3. Find some cases that match:
    ```{r}
    Matches <-
      Sample %>%
      filter(grepl("PO ", address))
    ```
    ```{r}
    Matches
    ```
4. Find cases that don't match:
    ```{r}
    Dont <-
      Sample %>%
      filter( ! grepl("PO ", address))
    ```
    ```{r}
    Dont 
    ```
5. Find any cases in the `Matches` that shouldn't be there.  They all look good here.
6. Find any cases in `Dont` that should have matched. (The `NCSU BOX` addresses should have matched our intension.)
7. It looks like `"BOX"` would have been a better pattern. We expect `"BOX"` to be followed by a space and a number, so `"\\s+BOX\\s+\\d+"` 
    ```{r}
    pattern <- "\\s+BOX\\s+\\d+"
    Matches <-
      Sample %>%
      filter(grepl(pattern, address))
    Dont <-
      Sample %>%
      filter( ! grepl(pattern, address))
    ```
    ```{r echo=FALSE, in_margin=TRUE}
    Dont
    ```
The result seems satisfactory.

Now we want to extract the box number. First, mark the part of the pattern that you want to extract by surrounding it with parentheses. Then, you'll be in a position to use `tidyr::extract()` to pull out the part of the pattern identified by extraction parentheses.

```{r}
pattern <- "\\s+BOX\\s+(\\d+)"
BoxNumbers <-
  Sample %>%
  filter(grepl(pattern, address)) %>%
  tidyr::extract(address, into="boxnum", regex=pattern)
```
```{r in_margin=TRUE}
BoxNumbers %>% head()
```

Note that `tidyr::extract()` should be given only those cases that match the regular expression, so `filter()` is applied before `tidyr::extract()`
