```{r include = FALSE}
library(DataComputing)
show_answers <- TRUE
```

Consider the following (pretty horrible) graphic. 

```{r q6plot, echo = FALSE, warning = FALSE}
CPS85 %>%
  ggplot(aes(x = educ, y = wage)) +
  geom_boxplot(aes(group=educ)) +
  geom_point(aes(color=sex)) +
  geom_smooth(method="lm", aes(color=sex)) +
  ylim(0,15) +
  labs(title = "Wage vs. Education in CPS85")
```

a. What are the glyphs/geoms in this graph? For each, say whether it reflects individual cases or reflects the collective properties of cases (a statistic).
```{block echo = show_answers}
    - `geom_point()` - individual
    - `geom_smooth()`- statistic
    - `geom_boxplot()` - statistic
```
b. What are the aesthetics of each glyph?
```{block echo = show_answers}
    - Points - x, y, color
    - Lines - x, y, color
    - boxplot -x, group,
```
c. Which variables are being mapped to each aesthetic?
```{block echo = show_answers}
    - x - Education
    - y - Wage
    - color - Sex
    - group - Education
```
#. Are the variables qualitative or quantitative?
```{block echo = show_answers}
    1. Education - quantitative
    1. Wage - quantitative
    1. Sex - qualitative
```
#. What are the guides on this graph?
```{block echo = show_answers}
    * labels/tickmarks on x and y axis showing the scale for education and wage
    * key on the side saying which color belongs to which gender
```
#. Reconstruct this graphic using `ggplot` and the `CPS85` data table.
```{r, echo = show_answers, eval = FALSE}
CPS85 %>%
  ggplot(aes(x = educ, y = wage)) +
  geom_boxplot(aes(group=educ)) +
  geom_point(aes(color=sex)) +
  geom_smooth(method="lm", aes(color=sex)) +
  ylim(0,15) +
  labs(title = "Wage vs. Education in CPS85")
```

Credit: Adam Lucas
