```{r include = FALSE}
library(DataComputing)
show_answers <- TRUE
```

The plot shows some relationships in a the `diamonds` data.

```{r include = FALSE}
set.seed(101)
diamonds <- 
  diamonds %>% 
  group_by(cut) %>%
  sample_n(size = 135)
```

```{r 153-diamonds, echo = FALSE}
diamonds %>% 
  ggplot(aes(x = carat, y = price, color = cut)) + 
  geom_point() +
  stat_smooth(color = "blue") + 
  facet_wrap( ~ cut)

```

The glyph-ready data are in this format:

```{r echo = FALSE}
diamonds %>% head()
```

1. What layers are on the graph? For each layer, say whether the glyph reflects a single case or the collective properties of cases (a statistic).
```{block echo = show_answers}
    Two layers:
    1. Glyphs are dots. Each dot represents a single case.
    2. A smoother through the dots. This is a statistic and shows collective properties of the cases.
```
2. What variables are being used in the graphic?
```{block echo = show_answers}
    `caret`, `price`, and `cut`
```
3. Which variables define the frame?
```{block echo = show_answers}
    `price` vs `carat`
```
4. Which variable or variables are being used to define facets?
```{block echo = show_answers}
`cut` defines the facets.
```
5. What are the aesthetics used? For each aesthetic, say which variable is being mapped to it or what constant it is being set to.
```{block echo = show_answers}
    - x: `carat`
    - y: `price`
    - color (of points): `cut`
    - color of smoother: set to blue
```
6. Assuming that the relevant packages are already loaded, write down a `ggplot()` command that will generate the graphic.
```{r echo = show_answers, ref.label="153-diamonds", fig.keep = "none"}
```
