```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(printr)
library(mosaic)
library(tidyr)
knitr::opts_chunk$set(tidy=FALSE, message=FALSE, warning=FALSE)
options(width = 80)
```

# Wide vs Narrow Data Tables

A data table is comprises *cases* and *variables*.

Each *variable* comprises *values* (or levels).

There is no hard distinction between a variable and a value.  What's a variable in one situation may be a value in another, and vice versa.

```{r echo=FALSE}
Students <-
  data.frame(who = c("Alice", "Lesley", "Yu"),
             x = c(7, 19, 23), 
             y = c("English", "Mandarin", "French"),
             dorm = c("Doty", "Doty", "Kirk"))
```

A data table

```{r}
Students
```

## Cases, Variables, and Values

* Variables: Who, X, and Y
    - Values: 
        - Who is a person's name
        - X is numeric
        - Y is a language name
        - dorm is a building name
* Cases: Alice, Lesley, Yu

```{r echo=FALSE}
Students
```

## Two formats

* Narrow

```{r echo=FALSE, warning=FALSE}
Students %>% 
  gather(key, value, x, y)
```

* Wide

```{r echo=FALSE, warning=FALSE}
Students
```

## Narrow and Wide

Data in Key/Value format are **narrow**

The corresponding **wide** format has 

* separate variables for each level in Key
* sets the values for those variables from the info in Value

Narrow:
```{r echo=FALSE, warning=FALSE}
head(Students,2)
```

Wide:
```{r echo=FALSE, warning=FALSE}
Students %>% 
  gather(key, value, x, y) %>% filter(who %in% c("Alice", "Lesley"))
```

## Narrow is relative

```{r echo=FALSE, warning=FALSE}
Students %>% 
  gather(key, value, x, y, dorm)
```

## Too narrow


```{r echo=FALSE, warning=FALSE}
Students %>% 
  gather(key, value, x, y, dorm, who)
```

There's nothing to identify a case!

## Gather --- from Wide to Narrow

Syntax:
```
WideInput %>% 
  gather(key_name, value_name, ...)
```
The `...` are the variables to be gathered together, e.g.
```{r warning=FALSE, echo=1}
StudentsNarrow <- Students %>% gather(key, value, x, y)
StudentsNarrow
```

## Cases in Narrow data

Aside from Key and Value, all the other variables identify the case.

The gathering makes multiple rows for each row in the wide form.  The variables **not** used for narrowing are copied into the new multiple cases.


## Spread --- from Narrow to Wide

Syntax:
```r
NarrowInput %>% spread(key, value)
```

Process:    

1. Group by **all** variables other than Key and Value
    These groups become the cases
2. Create new variables for each level in Key
3. Within each group, spread out the Values into the new variables.

```{r}
StudentsNarrow %>% spread(key, value)
```


