# Extracting matching parts of regular expressions

```{r include = FALSE}
library(DataComputing)
```


The `extractMatches()` data verb from the `DataComputing` package helps you to identify the specific characters that match in a regular expression.

Details:

1. Like all data verbs, `extractMatches()` takes a data table as input and produces a data table as output. That data table will have one or more additional columns with the extractions

2. syntax is `Data_table %>% extractMatches(regex, var,...)`. The `var` is the variable to be used in matching with the regex. The `...` provides a way for you to name the extracted segments.

3. To indicate that you want to extract the content that matches part of the regexp, wrap that part in parentheses.

4. When there is no match `extractMatches()` returns `NA`.

Example 1:  This regex looks for the letters "n" or "r" followed by a vowel. Both the specific letter and the specific vowel are extracted.

```{r}
BabyNames %>% 
  head() %>%
  extractMatches("([nr])([aeiouy])", name)
```


Example 2: What are the most common end vowels for Bible names?

To answer this question, you have to extract the last vowel from the name.  The `extractMatches()` transformation function can do this.

```{r}
BibleNames %>%
  extractMatches( "([aeiou])$", name, vowel=1) %>%
  group_by(vowel ) %>%
  summarise( total= n()) %>%
  arrange( vowel, desc(total) )
```

Example 2: What Bible Names start and end with a vowel, and what are those vowels?  In this example, we'll arrange for the first match to be called `beg_vowel` and the second to be called `end_vowel`.

```{r}
BabyNames %>%
  extractMatches( "^([AEIOU]).*([aeiou])$", name, beg_vowel=1, end_vowel=2) %>%
  head()
```


#Tasks for you

1. Which Bible names in `BibleNames` have the word "man" in the meaning (not first or last word)?

```{r,include=show_answers}
BibleNames %>%
  filter(grepl("[[:blank:]]+man[[:blank:]]+",meaning))
```

2. Grab everything but the last letter of `names` in `BabyNames` that end with a vowel (example: Grab Ev from Eva)

```{r,include=show_answers,warning=FALSE}
BabyNames %>%
  extractMatches(pattern="(.*)[aeiouy]$", name, root=1) %>%
  head()
```


3. What will this regex extract?

```{r, echo=FALSE, eval=FALSE}
BabyNames %>% 
  extractMatches("^[A-Z](.).(.*)[aeiou]$", name, early=1, late = 2) %>% 
  head(10)
```

