# Modeling library use


```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(printr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(digits=2)
```

<!-- Data provided by Katy Gabrio gabrio@macalester.edu -->

```{r echo = FALSE}
load("Data/Inventory.rda")
LOC <- read.csv("Data/LOC-Codes.csv",
                stringsAsFactors = FALSE)
Inventory <-
  Inventory %>%
  mutate(Last.Issued.Date = lubridate::ymd_hms(Last.Issued.Date))
```

The Macalester College library inventory, as reported on July 30, 2015, consists of `r nrow(Inventory)` items.  The large majority of these are books, but there are many other formats as well. The following types compose 99.9% of the library holdings. 

```{r}
Inventory %>%
  group_by(Material.Format) %>%
  tally() %>% 
  arrange(n) %>%
  filter(cumsum(n) > nrow(Inventory)/1000) %>%
  arrange(desc(n)) %>%
  as.data.frame(.)
```

A record is kept of how many times each item has been checked out since the start of record-keeping, and how many times it has been checked out in the last year.^[Judging from the dates in the records, record keeping for the data studied here started in January 2001.]  In addition, the date of the most recent check out is recorded.

```{r}
Inventory %>%
  group_by(Material.Format) %>%
  summarise(events_per_item_total = sum(Issued.Count, na.rm=TRUE)/n(),
            events_per_item_YTD = sum(Issued.Count.YTD, na.rm=TRUE)/n(), count=n()) %>%
  arrange(desc(events_per_item_total)) %>% 
  filter(count > 500) %>%
  as.data.frame()
```

Presumably, some kinds of items (e.g. newspapers) are not checked out.  Others are in obsolete formats (e.g. audiobook cassettes). By far the most heavily used items in the inventory are "Objects" and "otr."  For books and journals, by far the largest components of the library's collection, use is much less.  The average book has been checked out about 2 times since January 2001. Journals seem not to be checked out much.  This could be because of sparse use or because they are used without being checked out.

## Books

The analysis presented above indicates that the "average book" has been used about 2 times since 2001, and about 0.2 times in the last year.  It's important to note, however, that book use is highly skewed.  Some books are used many times; many books are not used at all. Among other things, this may depend on the subject of the book.


```{r}
CircBooks <-
  Inventory %>%
  filter(Material.Format == "Book", 
         grepl("Stacks - Level [2|3|4]$", Shelving.Location) |
         grepl("Service Desk - Res", Shelving.Location))
```


## Analysis

Write a summary of library book usage with an eye toward giving advice to the library administration on how many books can be archived or disposed of and an estimate of the yearly cost of providing alternative access though inter-library loan.

In order to make such an estimate, you'll need to estimate how much inter-library loan costs. Your figure does not need to be exactly accurate as you make the estimate. You can easily repeat the analysis with any new value.

As part of your presentation, tabulate results separately for each of the Library of Congress subject codes (in `LOC-CODES.csv`). You can access the data on inventory at  `http://tiny.cc/dcf/Inventory.rda`.  A very abbreviated list of book subjects, based on the Library of Congress number, is available at `http://tiny.cc/dcf/LOC-Codes.csv`.