# Files and Documents

## The "file path"

The "file path"" is the set of successive folders that bring you to the file.

There is a standard format for file paths. An example: `/Users/kaplan/Downloads/0021_001.pdf`. Here the filename is `0021_001.pdf`, the filename extension is `.pdf`, and the file itself is in the Downloads folder contained in the kaplan folder, which is in turn contained in the Users folder. The starting / means "on this computer".

## Using a file browser to find the path

Many people are used to finding files interactively with the mouse and may not be aware of the path to the directory holding the file.

The R `file.choose()` — which should be used only in the console, not in an Rmd file — brings up an interactive file browser. You can select a file with the browser. The returned value will be a quoted character string with the path and file name.

`file.choose()` then select a file. The output is:
`[1] "/Users/kaplan/Downloads/0021_001.pdf"`

For example I have a `.csv` file on my Desktop called `names.csv`. If I want to load it into R I might first find the path:

```{r eval = FALSE}
file.choose()
```
```{r echo = FALSE}
"/Users/Adam/Desktop/names.csv"
```

Then to load it I type:
```{r eval = FALSE}
my_names <- read.file("/Users/Adam/Desktop/names.csv")
```

Some common filename extensions for the sort of web resources you will be using:

* `.png` for pictures
* `.jpg` or `.jpeg` for photographs
* `.csv` or `.Rdata` for data files
* `.Rmd` for the human editable text of a document (called R Markdown)
* `.html` for web pages themselves

Credit: Adam Lucas
