# Trends in Popularity of Names

The relative popularity of different names for babies varies over the years and decades.  Let's construct a visualization of how the popularity of names varies in time.  


## Objective

Create a graph like the following using name of interest to you. 

```{r echo=FALSE,message=FALSE}
myNames <- data.frame( name=c("Abraham","Franklin",
                              "Jefferson","Washington", 
                              "Winston", "Adolf"), 
                       stringsAsFactors = FALSE )
Results <- 
  BabyNames %>%
  inner_join( myNames ) %>%
  group_by( name, year ) %>%
  summarise( total=sum(count) ) %>%
  mutate( total=100*total/sum(total))
```
```{r echo=FALSE,message=FALSE}  
Results %>% ggplot(aes(x=year,y=total,group=name)) +
  geom_line( size=1, alpha=.5, aes(color=name)) +
  ylab("Popularity") + xlab("Year")
```

The raw material you have is the `BabyNames` data set in the `DataComputing` package.

Point out which variables are categorical.  These can potentially be used for defining groups of cases.

### First, Individually ...

#### Step 1.

Analyze the graphic to figure out what a glyph-ready data table should look like. Mostly, this involves figuring out what variables are represented in the graph.  Write down a small example of a glyph-ready data frame that you think could be used to make something in the form of the graphic.

* What variable(s) from the raw data table do not appear at all in the graph?
* What variable(s) in the graph are similar to corresponding variables in the raw data table, but might have been transformed in some way.

#### Step 2

Consider how the cases differ between the raw input and the glyph-ready table.

* Have cases been **filtered** out? 
* Have cases been grouped and **aggregated/summarized** within groups in any way?
* Have any new variables been introduced? 
    
#### Step 3

Using English, write down a sequence of steps that will accomplish the transfiguration from the raw data table to your hypothesized glyph-ready data table.

#### Step 4: Confer with your colleagues

As a group, compare your different analyses in Steps 1 through 3.  Your goal is to develop a consensus for the design in Step 3.



#### Step 5: Implementation

Now you can start writing the commands themselves.  Do so, try to identify and solve any problems that arise, and make your glyph-ready data.

For graphing, you can use this template:
```{r eval=FALSE}
Results %>% ggplot(aes(x=year,y=total,group=name)) +
   geom_line( size=2, alpha=.5, aes(color=name)) +
   ylab("Popularity") + xlab("Year")
```

