# Example: Gender-neutral names: when wide is easier

```{r include = FALSE}
library(DataComputing)
```

Some operations are easy in wide format, but hard in narrow and *vice versa*

## Excerpt from `BabyNames`

```{r echo=FALSE}
data(BabyNames, package="DataComputing")
ShortBabyNames <- BabyNames %>%
  filter( name %in% c("Alice", "Lesley", "Yu"),
          year %in% 1998:1999 ) %>%
  arrange(name, year)
```

```{r echo=FALSE}
ShortBabyNames
```

Questions:

1. How many babies of each name and sex?
2. For each name, is it primarily given to girls or boys?  Which names are gender neutral?


## In narrow format

```{r echo=FALSE}
data(BabyNames, package="DataComputing")
BabyNames <- BabyNames %>%
  filter( name %in% c("Alice", "Lesley", "Yu"))
```

1. How many babies of each name and sex?

```{r}
BabyTotals <-
  BabyNames %>%
  group_by(name, sex) %>%
  summarise(total = sum(count))
```
```{r echo=FALSE}
BabyTotals
```

Easy!

## In Wide format

2. For each name, is it primarily given to girls or boys?  Which names are gender neutral?

- `sex` is the key variable
- `total` is the value variable


```{r echo=1:2}
BabyTotalsWide <- BabyTotals %>%
  spread(sex, total)
BabyTotalsWide
```

## With sexes side by side ...


```{r}
BabyTotalsWide %>%
  mutate(gender_specificity = pmax(F/(M+F), M/(M+F)))
```
