```{r include=FALSE}
require(mosaic)
require(knitr)
require(NHANES)
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
```

# Distributions

Often there is information to be found in the way cases differ one from the other.

A **distribution** is a representation of the patterns of spread of the cases.

## One variable

### Categorical: how many cases at each level?

Order: help the eye

### Quantitative

No discrete levels.  The fundamental display is **density**

Show some densities.

## Two or more variables: Displaying relationships

## A distribution template

### Aesthetics: 1, 2, or three variables

* `x = `  The first variable: *always!*
* `y = `  The *count* or *density* or another variable
* `group = ` or `fill = ` or `color = ` Still another variable

### Positions

* Dodged (the default)
* Stacked
* Proportion

### Facets 

One or two more categorical variables    

- `facet_wrap( ~ var)` 
- `facet_grid(var1 ~ var2)`

## Variable types

interval  ---- ordered  ----- categorical

1. Quantitative: a natural order and interval
2. Ordinal: a natural order but not an interval
3. Categorical: no natural order

## Order

The different levels of the variable have a natural order; it's natural to say that one level is higher than another or that a level is in-between two other levels.    

* Temperature
* Age
* Likert scale (strongly disagree, disagree, neutral, ...)
* Income bracket, e.g. 0-25K, 25K-40K, 40K-75K, 75K-150K
* Numerical scale, e.g. pain level of 1 to 10
* but not, e.g. Language, Nationality, ...

## Interval

For variables with a natural order, does the interval between levels have a specific meaning?   

* Temperature: Yes
* Age: Yes
* Likert scale: not really, but sometimes assumed
* Income bracket: No
* Language: No! (It's not even ordered!)

## NHANES

`NHANES` package has the `NHANES` data table


### Categorical variables:

* `Gender`: no order (therefore not interval)
* `Education`: ordered but not interval
* `SexOrientation`: not really ordered (therefore not interval)
* `Work`: not really ordered (therefore not interval)
* `MaritalStatus`: no order (therefore not interval)
* `HHIncome`: ordered but not interval

### Quantitative variables

* `UrineVol1` ordered and interval
* `Age`: ordered and interval

## Single variables

### Interval (hence ordered) --- use density

```{r}
NHANES %>%
  ggplot(aes(x = UrineVol1)) + 
  geom_density()
```

### Not interval --- use counts

But respect the order if there is one.

1. Quantitative: a natural order and interval
2. Ordinal: a natural order but not an interval
3. Categorical: no natural order

### Two variables

```{r}
NHANES %>% 
  ggplot(aes(x = Age)) +
  geom_density(aes(fill = MaritalStatus))
```

How could you make this better?

```{r}
NHANES %>%
  ggplot(aes(x = UrineVol1)) +
  geom_density(aes(group = MaritalStatus), 
               position = position_stack())
```

```{r}
NHANES %>%
  ggplot(aes(x = Age)) +
  geom_density(aes(fill = MaritalStatus), 
               position = position_stack())
```

```{r}
NHANES %>%
  ggplot(aes(x = Age)) +
  geom_density(aes(fill = MaritalStatus), 
               position = position_fill())
```

```{r}
NHANES %>%
  ggplot(aes(x = UrineVol1)) +
  geom_density(aes(fill = MaritalStatus, color = Gender), 
               position = position_stack())
```

```{r}
NHANES %>%
  ggplot(aes(x = UrineVol1)) +
  geom_density(aes(fill = MaritalStatus), 
               position = position_dodge())
```

```{r}
NHANES %>%
  ggplot(aes(x = Age, y = UrineVol1)) +
  geom_density2d()
```

```{r}
NHANES %>%
  ggplot(aes(x = Age, y = UrineVol1)) +
  geom_hex() +
  geom_smooth()

```
