# Regular expressions

## Visualizing matches

See the `str_view()` and `str_view_all()` widgets in [version 1.1 of stringr](https://blog.rstudio.org/2016/08/24/stringr-1-1-0/).


## Example

Pull out 100,000 names from `BabyNames`, adding up the totals over the years by sex.  Call this `NameList`.

```{r}
NameList <-
  BabyNames %>%
  group_by(name, sex) %>%
  summarise(total = sum(count))
```


* Names ending with "a".
* Names ending with a vowel.
* Names ending with a vowel or "y".
* Names with 3 consonants in a row.
* Names with 3 vowels in a row.

Interactive site for testing expressions: <http://regexone.com/>

## Extraction

With just one pattern to match, use `stringr::str_extract()`.

What are the most common vowels, by sex
```{r}
NameList %>%
  mutate(vowel = stringr::str_extract(name, "([aeiou]+)$")) %>%
  group_by(sex, vowel) %>%
  summarise( total = sum(total) ) %>%
  arrange(sex, desc(total)) %>%
  spread(key=sex, total)
```

With multiple patterns, use `tidyr::extract()`.
